\chapter{Project Plan}

\section{Project Objectives}
This project aims to produce a tool that notifies users of delays of the London buses they are waiting for, so that they can choose alternate routes and save waiting time.

\begin{description}
	\item[Back-end]  A timetable of bus journey times with an interface that supports real time query.
    \item[Front-end] A mobile application that allows users to check bus arrivals times with delay information at a give time and location. 
\end{description}

\subsection{Core Features}
    \begin{enumerate}
        \item Given user current location, query for predicted bus arrivals at the nearest bus stop.
 		\item Give user current location and destination, query for routes and suggest bus routes with delays taken into consideration.
        \item Given starting location, destination, desired departure / arrival time, suggest bus routes with delays taken into consideration.
    \end{enumerate}

\subsection{Optional Extensions}
    \begin{enumerate}
    	\item Optimise prediction accuracy 
        	\begin{itemize}
            	\item Analyse the average bus journey time for given starting and end stops to remove outliers. 
                \item Integrate iBus actual arrivals data to improve arrivals timetable precision.
            \end{itemize}
        \item Include other modes of transportation for alternate route recommendation.
        \item Gather more user feedback on mobile application and improve accordingly .
        \item Publish mobile application on android market.
    \end{enumerate}

\section{Current Progress}
\begin{enumerate}
  \item Obtained access to TFL Live bus arrival API stream.
  \item Designed databases schema.
  \item Setup databases to store bus arrival data.
  \item Setup remote server to run daemon process to pull data from live bus arrivals feed and store data into the database.
  \item Collected 2 weeks of bus arrivals data for analysis.
  \item Created sample bus arrivals timetable for a chosen pair of neighbouring bus stops (10002 - 11469).
\end{enumerate}

\section{Development Plan}
\begin{description}
	\item \textbf{Spring Term Week 4} 
    \begin{itemize}
    	\item Optimise arrivals table for fast query (e.g. via proper indexing).
		\item Create bus journey timetables with delay for all pairs of neighbouring bus stops.
    \end{itemize}
    \item \textbf{Spring Term Week 5}
    Develop journey planning feature with given start and end stops
    \item \textbf{Milestone 1 - Spring Term Week 6} 
    Design query to request journey time from bus stop A to bus stop B at a given time of the day.

	\item \textbf{Spring Term Week 7} 
    \begin{itemize}
    	\item Develop back-end API to receive query from mobile application. 
    	\item Start initial mobile interface design.
    \end{itemize}
	\item \textbf{Spring Term Week 8} 
	Design mobile application interface to extract user current location, and identify the nearest bus stop.
    \item \textbf{Milestone 2 - Spring Term Week 9} 
    Connect mobile application to send query to the back-end to request for bus arrival times for all routes reaching ther nearest bus stop. 
    \item \textbf{Spring Term Week 10}
    Improve mobile application user interface (gain user feedback and improve incrementally).
	\item \textbf{Fall-back Point - Easter Break} 
    \begin{itemize}
    	\item Implement mobile interfaces to get journey destination for journey planning. 
        \item Display suggested journey routes and travel time for a give time. 
    	\item Optimise delay predictions by filtering out anomalies in average journey time calculation
    	\item Evaluate performance by comparing with iBus data
    \end{itemize}
    \item \textbf{Milestone 3 - Summer Term Week 1 - 3} Implement extensions and start writing report.
	\item \textbf{Summer Term Week 4} Report first draft.
    \item \textbf{Summer Term Week 5} Report second draft.
    \item \textbf{Summer Term Week 6} Presentation first draft.
    \item \textbf{Summer Term Week 7} Presentation second draft.
    \item \textbf{Summer Term Week 8}
    	\begin{itemize}
        	\item \textbf{Tuesday, 16th June 4pm} Final report due.
    		\item Presentation practice.
        \end{itemize}
    \item \textbf{Summer Term Week 9, Monday 22nd and Tuesday 23rd June} Presentation days.
    \item \textbf{Summer Term Week 10, Monday 29th June} Final Archive due.
\end{description}